% GENERAL FORMAT OF EACH SLIDE
% ----------------------------

%\begin{frame}

%\newlibtitle or \implibtitle % for new vs. improved libraries

%\bookpath{path/to/the/library}: % directory or file, use macro for font
% Short description of what this library is, 1-2 lines.
%\begin{itemize}
%\item
% A highlight about this new or improved library.
%\item
% Another highlight.
%\item
% Say if it's described in a paper at this workshop,
% in which case, there may be little to say here.
%\end{itemize}

%\separation % to separate the next entry, if any

%\bookpath{another/path}: % as above

%\end{frame}

% ORDER OF THE SLIDES AND ENTRIES
% -------------------------------

% First all the new libraries, in alphabetical order.

% Then all the improved libraries, in alphabetical order.

% Then one or more slides with additional contributions.

% Don't worry about adding your contributions in order:
% Alessandro will quickly put them in order if needed,
% and also fix any formatting issues.

% TITLE SLIDE
% -----------

% Add your name as co-author, in alphabetical order of last name.

% Add your organization, in alphabetical order.
% Renumber the \inst{n} instances as needed
% (both inside \author{...} and inside \institute{...}).

% Or ask Alessandro to do this.

% OTHER REMARKS
% -------------

% Use the \code{...} macro for code (i.e.\ fixed-width) font within the text.

% If you want to show code block snippets,
% use the verbatim environment,
% and add [fragile] just after \begin{frame}.

% For guidelines on content and level of detail,
% please see existing entries.
% The goal is to provide enough information to get possible users interested.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{beamer}
\usetheme{Copenhagen}
\usecolortheme{whale}
\usepackage{inconsolata}

\beamertemplatenavigationsymbolsempty

\setbeamertemplate{footline}[frame number]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\code}[1]{\texttt{#1}}
\newenvironment{codeblock}{\begin{alltt}}{\end{alltt}}

\newcommand{\bookpath}[1]{\textbf{\code{#1}}}

\newcommand{\newlibtitle}{\frametitle{New Libraries}}
\newcommand{\implibtitle}{\frametitle{Improved Libraries}}

\newcommand{\separation}{\vspace*{1.5ex}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{What's New in the Community Books}

\subtitle{Since the ACL2-2022 Workshop}

\author{Alessandro Coglio\inst{1,5,6} (presenter),\\
        Warren Hunt\inst{3,8},
        Matt Kaufmann\inst{3},\\
        Ankit Kumar\inst{7},
        Vivek Ramanathan\inst{3},\\
        David Russinoff\inst{2},
        Eric W. Smith\inst{5,6,1},\\
        Sol Swords\inst{4},
        Max von Hippel\inst{7}}

\institute{\inst{1}Aleo Systems Inc.,
           \inst{2}{Arm Inc.},
           \inst{3}ForrestHunt Inc.,\\
           \inst{4}Intel Corp.,
           \inst{5}Kestrel Institute,
           \inst{6}Kestrel Technology LLC,\\
           \inst{7}Northeastern University,
           \inst{8}University of Texas at Austin}

\date{ACL2-2023 Workshop}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\frame[noframenumbering,plain]{\titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\frametitle{Overview}

\begin{itemize}
\item Almost 6,000 non-merge commits since the last Workshop.
\item From several contributors from several organizations.
\item Spanning hardware, mathematics, cryptography, blockchain,
      programming languages, virtual machines, machine code,
      standards, analysis, synthesis, and more.
\item These slides provide succinct summaries
      of major items in the book release notes,
      ordered by book path within each of the new and improved library parts.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\newlibtitle

\bookpath{kestrel/acl2data}:
Tools to collect machine learning data from the community books.
\begin{itemize}
\item Collected data is put into \code{*\_\_acl2data.out} files.
\item Data is collected by breaking proofs systematically
      and recording resulting information, including checkpoints.
\item More information in \code{README} and  \code{gather/README}.
\item Used by Proof Helpers in \code{kestrel/helpers}.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\newlibtitle

\bookpath{kestrel/big-data}:
Analyzing the ACL2 books as an object of scientific study.
\begin{itemize}
\item Work on identifying a maximal set of non-conflicting package definitions.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\newlibtitle

\bookpath{kestrel/built-ins}:
Tools to collect, document, and disable built-in events.
\begin{itemize}
\item Collect all the built-in event names in constants for
      \code{defun}s, \code{defthm}s, \code{defaxiom}s, etc.
\item Define XDOC topics for different portions of those events
      (by rule class, by data types, etc.).
\item Provide macros to disable certain sets of built-in events,
      e.g.\ all logic-mode functions, or all `default' rules.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\newlibtitle

\bookpath{kestrel/hints}:
Utilities for manipulating hints and parts of hints (combining, removing, renaming).
\begin{itemize}
\item Collected and improved existing utilities, and added new ones.
\item New \code{:casesx} hint to make all $2^n$ combinations of $n$ boolean expressions.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\newlibtitle

\bookpath{kestrel/tests}:
Tests of various constructs, to understand precisely what they do in all cases.
\begin{itemize}
\item Add tests of \code{b*}, \code{case-match}, \code{mv-let}, \code{defthm}, \code{defrule}, \code{xargs}, and \code{hints}.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\newlibtitle

\bookpath{kestrel/world-light}:
Lightweight world utilities.
\begin{itemize}
\item Querying the world, gathering event names, etc.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\newlibtitle

\bookpath{kestrel/zip}:
Handling zip files.
\begin{itemize}
\item Can unzip a file within ACL2.
\item Only handles Deflate compression (or no compression).
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\newlibtitle

\bookpath{projects/vwsim}:
Super-conducting circuit simulator system.
\begin{itemize}
\item Presented at the previous Workshop, now available for general use.
\item Simulator of RSFQ (Rapid Single Flux Quantum),
      a digital electronic device that uses superconducting devices
      to process digital signals.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\newlibtitle

\bookpath{workshops/2023/kumar-etal}:
Formal model of GossipSub protocol with various attacks.
(Discussed in a paper at this Workshop.)
\begin{itemize}
\item Model of entire protocol with special attention paid to score function.
\item Proof that the protocol satisfies some security properties but not others.
\item Examples of automated attack discovery using CGen.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\newlibtitle

\bookpath{workshops/2023/vonhippel-etal}:
Formal model of Retransmission TimeOut (RTO) computation.
(Discussed in a paper at this Workshop.)
\begin{itemize}
\item Proofs in ACL2, ACL2s, and ACL2(r) that for all $\alpha \in [0, 1)$,
      $\lim_{n\to\infty}\alpha^n = 0$.
\item Multiple proof strategies in ACL2s.
\item Formal model of RTO system and proven asymptotic bounds on its variables.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{build}:
\code{cert.pl} build system.
\begin{itemize}
\item New support for builds that involve saved images.
\item Support for \code{ACL2\_PROJECTS} file specifying project directories for book relocation.
\item \texttt{build/include-events}: utility for including books that
  are bundles of other books without needing to certify the bundles.
\item For Jenkins: improve \code{Makefile}, improve script printing and documentation, and limit the max load.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{centaur/sv} and \bookpath{centaur/vl}:
Centaur/Intel hardware verif framework.
\begin{itemize}
\item New methodology for datapath proofs with cutpoint decomposition.
\item New dual-run comparison mode for SVTV-CHASE hardware simulation debugging tool.
\item Many bugfixes and small improvements.
\end{itemize}

\bookpath{centaur/fgl}:
Centaur/Intel bitblasting rewriter.
\begin{itemize}
\item New capability to solve conjectures with 37 or fewer Boolean variables by exhaustive testing.
\item Many small improvements and added features.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{demos}:
Demos of ACL2 and tools built on ACL2.
\begin{itemize}
\item Added examples of \code{include-raw}.
\item Added examples of \code{partial-encapsulate}.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{doc}:
ACL2 + Books Documentation.
\begin{itemize}
\item Added Top 100 Theorems list for ACL2.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/abnf}:
Augmented Backus-Naur Form (ABNF).
\begin{itemize}
\item Refactored to organize constituents more clearly.
\item Significantly extended the parsing generation tools.
\item Added new tools for ingesting grammars
      and generating operations and theorems for them.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/acl2-arrays}:
Reasoning about ACL2 arrays (\code{aref1}, \code{aset1}, etc.).
\begin{itemize}
\item Added and improved rules.
\item Improved organization and documentation.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/alists-light}:
Lightwight library about alists.
\begin{itemize}
\item Added and improved rules.
\item Added \code{string-string-alistp}, \code{clear-keys}.
\item New book on \code{symbol-symbol-alistp}.
\item Deprecated \code{lookup-equal-lst} (use \code{map-lookup-equal}).
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/arithmetic-light}:
Lightwight library about arithmetic.
\begin{itemize}
\item Added and improved rules.
\item Added \code{log2} (floor of log, for any positive rational).
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/arrays-2d}:
Two-dimensional arrays as lists of lists.
\begin{itemize}
\item Various improvements (some suggested by \code{improve-book} tool).
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/apt}:
Automated Program Transformations (APT).
\begin{itemize}
\item Improved the \code{simplify}, \code{wrap-output},
  \code{drop-irrelevant-params}, \code{rename-params}, and
  \code{finite-difference} transformations.
\item Improved the robustness of the proofs
  generated by the \code{restrict} transformation.
\item Added the \code{def} utility for calling transformations.
\item New transformation \code{rename-calls}.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/axe (1)}:
The Axe Toolkit.
\begin{itemize}
\item Many improvements and fixes (clarified code, added and verified guards, removed skip-proofs from legacy tools, new proofs about implementation, deprecated old code, fixed names, more robust proofs).
\item Added, improved, and organized rules.
\item Improved rule-lists and built in more rules.
\item Improved printing, error handling, and reporting.
\item Improved counterexample machinery.
\item Optimized: Use hash-tables and fast-alists, avoid some quadratic behaviors, use \code{tshell-call} instead of \code{sys-call}.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/axe (2)}:
The Axe Toolkit.
\begin{itemize}
\item Improved pruning (apply to DAGs, use more assumptions, evaluate new ground terms) and use during lifting.
\item Evaluate more ground terms.  Improve use of memoization.
\item The \code{ACL2\_STP\_VARIETY} environment variable now governs how STP is to be called (which option syntax).
\item Proved some theorems validating Axe's translation to SMT solvers.
\item Improved \code{unroll-spec-basic} (set of rules used).
\item Added tests.  Improved documentation.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/axe (3)}:
The Axe Toolkit.
\begin{itemize}
\item Added \code{:max-conflicts} option to query tool.
\item Renamed \code{prove-equivalence2} to \code{prove-equal-with-tactics}.
\item Allow each generated prover to have its own set of \code{default-global-rules}, add \code{:extra-global-rules} option.
\item Fix rare completeness issue in provers.
\item Add \code{:var-ordering} option to prevent undesirable substitutions.
\item Add new prover tactic \code{:rewrite-top}.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/axe/jvm}:
The Axe Toolkit for JVM.
\begin{itemize}
\item Added JVM Formal Unit Tester.
\item Simplified \code{unroll-java-code} in several ways.
\item Added examples: Axe proofs of AES implementations, Formal Unit Tester examples.
\end{itemize}

\bookpath{kestrel/axe/r1cs}:
The Axe Toolkit for R1CSes
\begin{itemize}
\item Built in more global rules.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/axe/x86}:
The Axe Toolkit for x86.
\begin{itemize}
\item Added prototype Formal Unit Tester for x86 binaries.
\item Added support for parsing and lifting ELF files.
\item Misc improvements to x86 lifting.
\item More readable normal forms, better debugging.
\item Added \code{:position-independent} option.
\item Added examples: TEA block cipher, recursive factorial, popcount.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/booleans}:
Rules and definitions involving booleans (supports Axe).
\begin{itemize}
\item New and improved rules.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/bv}:
Bit vectors.
\begin{itemize}
\item Added, improved, renamed, and organized rules.  Improved names.
\item Macro fixes.  Tweaked \code{repeatbit}.
\item New scheme for converting terms to use bv functions.
\item Added \code{bitxor\$}, \code{bitand\$}, and \code{bitor\$} (faster, stricter guards).
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/bv-lists}:
Lists of bit vectors.
\begin{itemize}
\item Added and organized rules.
\item New bv-array conversion utilities.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/c}:
Models, proofs, and tools for C.
\begin{itemize}
\item Refactored to organize constituents
      (deep embedding, shallow embedding, code generator)
      more clearly.
\item Extensions and improvements to the model of C.
\item Significant extensions of the C code generation capabilities.
\item Significant performance improvements of the generated correctness proofs,
      via a new modular proof approach.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/clause-processors}:
Clause processors and utilities for them.
\begin{itemize}
\item Organize library, misc cleanups.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/crypto}:
Cryptography.
\begin{itemize}
\item Added a formal spec for the AES block cipher.
\item Added a formal spec of the SHA-3 hash function (proved equivalence to Keccak spec in some cases).
\item Added a formal spec of the TEA block cipher.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/crypto/pfcs}:
Prime Field Constraint Systems (PFCS).
(Discussed in a paper at this Workshop.)
\begin{itemize}
\item Added a concrete syntax.
\item Improved the abstract syntax.
\item Added many theorems,
      including proof support rules for compositional reasoning.
\item Added a translator from R1CS to PFCS,
      along with a checker for the R1CS subset of PFCS.
\item Added a proof-generating lifter from deeply to shallowly embedded PFCS.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/crypto/r1cs}:
Rank-1 constraint systems.
\begin{itemize}
\item Added and generalized rules.
\item Added dense form of R1CSes (legacy).
\item Added optimized function \code{r1cs-constraint-list-vars}.
\item Imported \code{primep} into package.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/evaluators}:
Evaluators (e.g., to support meta-reasoning).
\begin{itemize}
\item New tool \code{defevaluator-theorems}.
\item Added theorems.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/file-io-light}:
Lightwight library about file input and output.
\begin{itemize}
\item Added rules about \code{princ\$}, \code{prin1\$}, \code{open-output-channel}, \code{open-output-channel!}, etc.
\item New books on \code{close-input-channel}, \code{close-output-channel}, \code{typed-io-listp}, \code{print-object\$-fn}, and \code{newline}.
\item Added \code{read-objects-from-file-with-pkg}, \code{read-file-into-line-list}, and \code{read-file-into-line-list-no-error}.
\item Added and improved rules.  Organized and optimized.  Fixed names.
\item Proved that various functions don't change the world.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/floats}:
Cleanroom formalization of IEEE-754 floating point (in progress).
\begin{itemize}
\item Model of floating point formats, representable rationals, encoding/decoding, normals, subnormals, zeros, infinities, NaNs, comparisons.  No rounding yet!
\item Tries to match the official spec as closely as possible.
\item Proved some connection theorems to the RTL formalization.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/helpers}:
Helper tools for finding / repairing proofs and improving ACL2 books.
\begin{itemize}
\item New Proof Advice tool to get proof help over the web, based on ML models and other heuristics.
\item Tools to evaluate proof advice on one or many books.
\item Improvements to \code{improve-book} tool.  Improve linter and move it here.
\item New \code{speed-up-event} utility that tries to speed up theorems.
\item Add initial \code{repair} tool to fix broken proofs.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/htclient}:
Communication via HTTP.
\begin{itemize}
\item Lightweight variant of HTTP POST.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/json-parser}:
JSON Parser.
\begin{itemize}
\item Added \code{parse-string-as-json}, \code{defconst-from-json-file}.
\item Improved implementation (more tail recursion).
\item Added and improved rules.
\item Organized and sped up proofs.
\item Added and improved tests.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/jvm}:
JVM model.
\begin{itemize}
\item Clarified model and added new rules.
\item Better normal forms and abstractions.
\item Renamed \code{load-class-XXX} to \code{read-class-XXX}.
\item Support for reading code directly from jar files.
\item Put more names in \code{JVM} package.
\item Worked on verifying guards.
\item Improved handling of floats/doubles, LDC bytecode.
\item Improved error reporting in class file parser.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/lists-light}:
Lightweight lists library.
\begin{itemize}
\item Added various rules, organized library.
\item New books on \code{position-equal-ac}, \code{position-equal}.
\item Added \code{filter-non-members}.
\item Added \code{union-equal-alt} (keeps earlier duplicates).
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/meta}:
Metafunctions / meta rules.
\begin{itemize}
\item Add some experiments verifying rewriter-like tools using \code{meta-extract}.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/number-theory}:
Number theory library.
\begin{itemize}
\item Import some symbols into the \code{PRIMES} package, including \code{dm}::primep.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/prime-fields}:
Prime field library.
\begin{itemize}
\item Add and improve rules, and simplify.
\item Import some symbols into the \code{PRIMES} package, including \code{dm}::primep.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/std/system}:
Standard system library (extension of Std).
\begin{itemize}
\item Added utilities
      \code{untranslate\$},
      \code{genvar\$},
      \code{one-way-unify\$}, and
      \code{termination-theorem\$},
      which are logic-mode variants of built-in utilities,
      via \code{magic-ev-fncall}.
\item Added utilities
      \code{guard-theorem-no-simplify} (program-mode) and
      \code{guard-theorem-no-simplify\$} (logic-mode),
      which are variants of \code{guard-theorem} that does no simplification.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/std/util}:
Standard system library (extension of Std).
\begin{itemize}
\item New utilities for error-value tuples have been added,
      to facilitate the generation, propagation, and catching of errors
      in statically strongly typed code.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/string-light}:
Lightwight library about strings.
\begin{itemize}
\item Added \code{string-starts-withp}, \code{add-prefix-to-strings},
  \code{split-string-last}, \code{strip-prefix-from-string},
  \code{strip-suffix-from-string}, \code{strip-suffix-from-strings},
  \code{decimal-digit-to-char},
  \code{decimal-digit-to-string},\code{strnthcdr},
\item New books about \item{subseq}, \code{strcar}, \code{strcdr}, \code{char}.
\item Renamed \code{string-ends-inp} to \code{string-ends-withp}
\item Added rules and tests.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/terms-light}:
Lightweight library for manipulating terms.
\begin{itemize}
\item New and improved rules. Various fixes. Generalized, optimized, and clarified.
\item New books about \code{all-fnnames1}, \code{trivial-formals}, \code{get-conjuncts}, and \code{get-hyps-and-conc}, and \code{replace-corresponding-arg}.
\item Improved \code{reconstruct-lets-in-term} to handle \code{mv-let} and ignored vals.
\item Improved \code{substitute-unnecessary-lambda-vars}.
\item Added \code{subst-var-alt} (avoids creating unserialized lambdas).
\item Proved that some utilities return / preserve closed lambdas.
\item Added tests.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/typed-lists-light}:
Lightweight library for lists of elements of various types.
\begin{itemize}
\item Added rules and organized library.
\item New books about \code{alist-listp}, \code{symbol-alist-listp}, \code{symbol-term-alist-listp}, \code{map-strip-cdrs}, \code{all-digit-charsp}.
\item Added \code{string-list-listp}.

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/untranslated-terms}:
Utilities about untranslated terms.
\begin{itemize}
\item Various improvements, especially about \code{case}, \code{case-match}, and \code{cond}.
\item Improvements for \code{b*} (e.g., support for \code{\&}, \code{list}, and \code{er} binders).
\item Added utilities for extracting conjuncts and disjuncts, and for getting free variables from terms.
\item Improved support for \code{let} and \code{mv-let}.
\end{itemize}

\bookpath{kestrel/untranslated-terms-old}:
Old utilities about untranslated terms.
\begin{itemize}
\item Moved to \code{kestrel/untranslated-terms/untranslated-terms-old}.
\item Fixed some checks.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/utilities (1)}:
Various utilities in the Kestrel books.
\begin{itemize}
\item Added \code{show-books} utility,
      which returns a tree representing
      the books included in the current ACL2 session.
\item Improved \code{checkpoint-list} and related utilities.
\item Added parser for CSV files and utils for handling parsed CSV files.
\item Split \code{byte-array-stobj} into separate book.
\item Proved that \code{merge-sort-symbol<} returns a sorted list.
\item Deprecated \code{defopeners-mut-rec} (use \code{defopeners}).
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/utilities (2)}:
Various utilities in the Kestrel books.
\begin{itemize}
\item New books on \code{read-acl2-oracle}, \code{update-acl2-oracle}, \code{invariant-risk}, \code{getenv\$}, \code{get-serialize-character}, \code{plist-worldp}, \code{char-code}, widening margins, \code{strip-cadrs}, \code{>=-len}, \code{our-digit-char-p}, \code{read-run-time}, \code{get-cpu-time}.
\item Added misc rules (e.g., about \code{world}, \code{state}, \code{channels}, \code{assoc-keyword}, setting margins, \code{coerce}, \code{non-trivial-bindings}, \code{acl2-count}).
\item New utilities: \code{most-recent-failed-theorem-goal}, \code{translatable-term-listp}, parsing of strings, \code{dir-of-path}, \code{all-included-books}, \code{defthms-in-world}, \code{sys-call-event}, \code{split-path}, \code{reduce-print-level}, \code{pack-in-package}, \code{eval-tests}, \code{fast-alist-set}, \code{array-stobj}, \code{setenv\$-event}, \code{prove\$-nice-trying-hints}, \code{prove\$-nice-with-time-and-steps}, \code{last-prover-steps\$}, \code{merge-sort-string<}.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/utilities (3)}:
Various utilities in the Kestrel books.
\begin{itemize}
\item Improved \code{prove\$-nice} and \code{prove\$+}, \code{declare} utilities, \code{defmergsort}, \code{defxdoc-for-macro}, utilities for temp dirs, and material about conjuncts/disjuncts and \code{acl2-count}.
\item New tests. Better error reporting.
\item Improved utilities: \code{defmergesort} (\code{:extra-theorems} option), \code{ld-history} utils, \code{ubi}, \code{defstobj+} (generate more theorems, e.g., about hash-table fields).
\item New utils \code{shuffle-array-stobj} / \code{shuffle-array-stobj2} and \code{shuffle-list} / \code{shuffle-list2}, using a Fisher-Yates shuffle.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/utilities/digits-any-base}:
Number representations in arbitrary bases.
\begin{itemize}
\item Added several theorems.
\item Extended the \code{defdigits} macro to generate more theorems.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/utilities/omaps}:
Ordered maps (omaps).
\begin{itemize}
\item Added several theorems.
\item Improved documentation and organization.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/x86}:
Kestrel's additions to support working with the \code{x86isa} model
\begin{itemize}
\item Improved and added various rules, e.g., about condition codes.
\item Clarified and improved rule lists (used by Axe).
\item Improved parsers, fixed ELF parser bug.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{kestrel/zcash}:
The Zcash protocol.
\begin{itemize}
\item Add \code{:var-ordering} option to \code{verify-zcash-r1cs}.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{projects/curve25519} and \bookpath{projects/shnf}:
Older developments by Russinoff.
\begin{itemize}
\item Switched to DM (discrete math) package.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{projects/group} and \bookpath{projects/numbers}:
Formalizations of group theory and number theory.
(Discussed in two papers and a rump talk at this workshop).
\begin{itemize}
\item Elementary finite group theory,
  essentially the content of an advanced undergraduate course.
\item Various results in elementary number theory.
\item \code{projects/numbers} was previously called
  \code{projects/quadratic-reciprocity}, and it has been moved to the DM
  package.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{rtl}:
Register Transfer Logic.
\begin{itemize}
\item Switch some utilities, such as \code{local-defun}, to ACL2 package.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{std/io}:
Standard I/O library.
\begin{itemize}
\item Extended \code{read-string}
  to take an optional package as argument.
\item Added a lightweight version of \code{read-string}.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{std/lists}:
Standard Lists library.
\begin{itemize}
\item Have \code{deflist} add \code{-compound-recognizer} to some rule names, to avoid clashes.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{std/util}:
Standard utilities library.
\begin{itemize}
\item The \code{er} binder of \code{b*} has been extended with
  an option \code{:iferr} to return an alternative value in case of error.
\item Calls of \code{defaggregate} can now (usually) be local, and its proofs are more robust.
\item Calls of \code{define} no longer turn on error output if it was off before the \code{define}.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{tools}:
Various tools.
\begin{itemize}
\item Substantially enhanced \code{with-supporters}.
\item Improved \code{prove\$}, e.g.\ to treat hard errors as ordinary failures.
\item Made \code{run-script} more robust against time variations.
\item Added tools \code{functions-after} and \code{macros-after},
  which return names of functions and macros introduced after a given name.
\item Added new \code{tools/top} book to gather tools (for inclusion in the
  manual, instead of individual tool books).
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{workshops/2022/russinoff-calendar}:
Modeling the Hebrew calendar.
\begin{itemize}
\item Switched to ACL2 package.
\end{itemize}

\bookpath{workshops/2022/russinoff-groups}:
Group theory.
\begin{itemize}
\item Switched to DM (discrete math) package.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\implibtitle

\bookpath{xdoc}:
XDOC documentation.
\begin{itemize}
\item Added support for Greek letters and more mathematical symbols.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
